Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-27 Manuscript/manuscript.Rmd
results edits

1. **Overlap Counts (OC)**: For each genomic bin, the total number of instances where a bin overlapped with regions assigned to genomic annotations (features).

@@

To identify the combination of data resolution, predictor type, and re-sampling technique necessary for optimal cell line specific TAD boundary prediction, we created an ensemble framework to systematically compare performances of different model inputs (Figure 2). The framework uses data for a particular cell line *L = {GM12878, K562}*, at a given resolution *R = {5 kb, 10 kb, 25 kb, 50 kb, 100 kb}*, using one of the predictor type *P = {OC, OP, Distance}*, and the re-sampling technique *B = {None, RUS, ROS, SMOTE}*. Once the model inputs were established, a random forest classifier was trained on chromosome-specific datasets using 10-fold cross validation. The model was then validated on the same separate chromosome specific test set for unbiased comparisons [??? Ugly sentence, rewrite] and evaluated using Matthews Correlation Coefficient (MCC). Letting $|X|$ denote the cardinality of each set, we evaluated $|L| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $2 \times 5 \times 3 \times 4 \times 21 = 2520$) models. Additional performance metrics are reported in the Supplementary materials.

![**A diagram of the ensemble framework detailing the combination of inputs used in the model building pipeline.** [??? Describe abbreviations on the figure, the figure itself is already described in text] ](Figures/figure3_ensemble_framework_V3.png)

>>>>

1. **Overlap Count (OC)**: For each genomic bin, the total number of instances where a bin overlapped with regions assigned to genomic annotations (features).

@@

To identify the combination of data resolution, predictor type, and re-sampling technique necessary for optimal cell line specific TAD boundary prediction, we created an ensemble framework to systematically compare performances of different model inputs (Figure 2). The framework uses data for a particular cell line *L = {GM12878, K562}*, at a given resolution *R = {5 kb, 10 kb, 25 kb, 50 kb, 100 kb}*, using one of the predictor type *P = {OC, OP, Distance}*, and the re-sampling technique *B = {None, RUS, ROS, SMOTE}*. Once the model inputs were established, a random forest classifier was trained on chromosome-specific datasets using 10-fold cross validation. The different combinations of inputs were evaluated by assessing the Matthews Correlation Coefficient (MCC) of each model [??? Ugly sentence, rewrite; !!!fixed]. Letting $|X|$ denote the cardinality of each set, we evaluated $|L| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $2 \times 5 \times 3 \times 4 \times 21 = 2520$) models. Additional performance metrics are reported in the Supplementary materials.

![**A diagram of the ensemble framework detailing the combination of inputs used in the model building pipeline.** Each cell line (GM12878, K562) specific model was composed of a combination of different data level characteristics including resolution (5 kb, 10 kb, 25 kb, 50 kb, 100 kb), predictor type (overlap count (OC), overlap percent (OP), $log_{2}$ distance (Distance)), and re-sampling technique (no re-sampling (None), random over-sampling (ROS), random under-sampling (RUS), synthetic minority over-sampling technique (SMOTE)). The following sets of genomic annotations were used to create each predictor type including histone modification (HM), chromating state segmentations (ChromHMM), transcription factor binding sites (TFBS). [??? Describe abbreviations on the figure, the figure itself is already described in text;!!!fixed] ](Figures/figure3_ensemble_framework_V3.png)
